The hydra and the rooster

### 2020-07-30 latex/chapter-ordinals.tex
cleanup Schutte/Countable; replace omega tactic with lia

Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image},
 \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
relation \texttt{wf} is well founded.

@@

This notion is defined generically as a class in
 Module~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic} see Fig~\vref{figure:subsegment}).

\begin{Coqsrc}
(** The segment associated with nA is isomorphic to
    the interval [0,b) *)

Class  Subsegment 
       `(nA : @OrdinalNotation A ltA stoA compareA)
       `(nB : @OrdinalNotation B ltB stoB compareB)
       (b :  B)
       (iota : A -> B):=
  {
  subseg_compare :forall x y : A,  compareB (iota x) (iota y) =
                                 compareA x y;
  subseg_incl : forall x, ltB (iota x) b;
  subseg_onto : forall y, ltB y b  -> exists x:A, iota x = y}.
\end{Coqsrc}

@@

   Global Instance F_incl_ij  : SubSegment  (FinOrd i) (FinOrd j) b iota_ij.
  (* ... *)

@@

@@

\begin{itemize}
\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap:T1}).
\item $\Gamma_0$: The set of terms in Veblen normal form.
\end{itemize}

>>>>

Applying lemmas of the libraries \texttt{Coq.Wellfounded.Inverse\_Image}, \linebreak
 \texttt{Coq.Wellfounded.Inclusion}, and \texttt{Coq.Arith.Wf\_nat}, we prove that our
relation \texttt{wf} is well founded.

@@

@@

   Global Instance F_incl_ij  : SubON  (FinOrd i) (FinOrd j) b iota_ij.
  (* ... *)

@@

\index{Projects}
\begin{project}
Define in \coq{} the product of two ordinal notations $N_A$ and $N_B$.
If $A$ [resp. $B$] is the underlying type of $N_A$ [resp. $N_B$], the
product \texttt{ON\_mult $N_A$ $N_B$} is implemented over the cartesian product $B\times A$ (with the lexicographic ordering).

For instance, the
elements of the product \texttt{ON\_mult Omega (FinOrd 3)} are ordered as follows.
\[(0,0),(0,1),(0,2),(0,3),(0,4),\dots,{\color{red}(1,0),} (1,1),(1,2),\dots, {\color{red}(2,0)},(2,1),(2,2),\dots\]

Note that the elements of  \texttt{ON\_mult (FinOrd 3) Omega} are differently ordered (without limit ordinals):
\[(0,0),(1,0),(2,0),(0,1),(1,1),(2,1),(0,2),(1,2),(2,2),(0,3),\dots\]
(no limit ordinals).

Prove that \texttt{ON\_mult (FinOrd $i$) Omega} is isomorphic to
\texttt{Omega}  whilst
\texttt{Omega}  is a sub-notation of \texttt{ON\_mult Omega (FinOrd $i$)},
for any strictly positive $i$.
\end{project}

@@

\section*{Other ordinal notations}

\index{Projects}
\begin{project}
Let $N_A$ be a notation system for ordinals strictly less than $\alpha$, 
with the strict order $(A,<_A)$. Please build the notation system
\texttt{ON\_Expl $N_A$}, on the type of multisets of elements of $A$
(or, if preferred, the type of non-increasing finite sequences on $A$,
provided with the lexicographic ordering on lists).

For instance, let us take $N_A=\texttt{Omega}$, and take $\alpha=(4,4,2,1,0)$,
 $\beta=(4,3,3,3,3,3,2)$, and $\gamma=(5)$. Then $\beta<\alpha<\gamma$.

In contrast the list $(5,6,3,3)$ is not non-increasing (\emph{i.e.} sorted w.r.t. $\geq$), so it is not to be considered.

Note that if the notation $N_A$ implements the ordinal 
$\alpha$,  the new notation $\omega^{N_A}$ must implement the ordinal $\varphi_0(\alpha)$, a.k.a. $\omega^\alpha$ (see chapter~\ref{chap:schutte})

\end{project}

\begin{remark}
 The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
in Veblen normal form (see 
\href{../src/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
\end{remark}
